---
title: "Data Pipeline"
author: "Oluwasegun Amoniyan"
date: "03/27/2024"
output: 
  github_document: 
    toc: TRUE
---

```{r setup}
##Set knitr options (show both code and output, show output w/o leading #)
knitr::opts_chunk$set(echo = TRUE, include = TRUE, comment=NA)

#load tidyverse
library(tidyverse)

#load excel
library(readxl)
```

# Data summary

International Corpus of English (ICE) Nigeria (ICE-Nig) corpus comprises 609,581 spoken words and 400,796 written words. The corpus hosts `text files` and `audio files` for different discourse, including `broadcast news`, `broadcat talks`, `classroom conversations`, `non-broadcast` shows, `interviews`, `phone calls`, `conversations`, and `demonstrations`. The website has a metadata Excel [file](https://sourceforge.net/projects/ice-nigeria/files/) that describes the background (`age`, `gender`, `ethnicity`, `profession`, type of `speech style`) of the participants. Only `broadcast news` and `broadcast talks` are sampled for the data wrangling and analysis because of the time frame. 
Specifically, attention will be paid to different realizations for NURSE vowels across different `profession`, `gender`, `age`, and `ethnicity` in the sampled files.


# loading raw files

The first file (as nurse_vowel) includes **nurse_vowel** that contains **file_name**, **formant trajectories**, **formant bindwidth**, **pitch**, **intensity** and **harmonity**
The second file (as nurse_segment_info) includes *segment information** 

The second file estimates the aggregate formant means for the winning acoustic features in the predictions at every 2ms for 20 times per NURSE vowel realization.

The third file provide segment info such as `context`, `segment duration`...
```{r}
#load raw file
nurse_vowel <- read_excel("C:/Users/oamon/Oluwasegun DS Project/Sociophonetic-study-of-NURSE-vowels-in-NE/untidy_excel/nurse_vowel.xlsx")

nurse_raw_aggregated <- read_excel("C:/Users/oamon/Oluwasegun DS Project/Sociophonetic-study-of-NURSE-vowels-in-NE/untidy_excel/nurse_raw_aggregated.xlsx")

unstructured_seg_info <- read_csv("C:/Users/oamon/Oluwasegun DS Project/Sociophonetic-study-of-NURSE-vowels-in-NE/untidy_excel/unstructured_seg_info.csv")
```


#rename the first column in `nurse_raw_aggregated`before joining the dataframes together.

# Working on `nurse_segment_info`
Data cleaning
  **remove** uninformative columns including `omit`, `previous_sound` and `next_sound`

```{r}
nurse_raw_aggregated <- 
  nurse_raw_aggregated %>%
 rename(
    file_name = contains("þÿfile"))
head(nurse_raw_aggregated)

nurse_raw_aggregated <- nurse_raw_aggregated%>%
mutate(files = map_chr(file_name, ~ str_sub(.x, 1, 7)))
head(nurse_aggregate)
```

# Tidy unstructed_seg_info

`unstructed_seg_info` has segment features for NURSE vowels.  
```{r}
nurse_segment_info <- separate(unstructured_seg_info, inputfile.outputfile.vowel.interval.duration.start.end.previous_sound.next_sound.omit.word.word_interval.word_start.word_end.previous_word.next_word, 
                into = c("inputfile", "outputfile", "vowel", "interval", "duration", "start", "end", "previous_sound", "next_sound", "omit", "word", "word_interval", "word_start", "word_end", "previous_word", "next_word"), 
                sep = ",")
```


```{r}
#before joining nurse_segment_info with the aggregate by outputfile, there must be a similar name between the two dataframes 
nurse_segment_info <- 
  nurse_segment_info %>%
 rename(
    file_name = contains("outputfile"))
head(nurse_segment_info)
```


**Check different NURSE vowel production by audio files**

```{r}
nurse_segment_info %>%
  select(inputfile)%>%
  table()
```



# remove uninformative columns in `nurse_segment_info`.

The following columns(`previous_sound`, `next_sound` and `omit`) are dropped in the `nurse_segment_info` because they add no value to the `dataframe`.


```{r}
nurse_segment_info <- nurse_segment_info %>%
  select(-c("previous_sound", "next_sound", "omit"))
```

**Hurray** `nurse_segment_info` is now tidy

# NURSE vowel realizations in NE

```{r NURSE vowels in NE}
nurse_segment_info%>%
  select(vowel)%>%
  table()
```

# Combine files
combine both files as a single `csv` file before `data analysis`

```{r}
#combined_data <- merge(nurse_segment_info, nurse_raw_aggregated, by = "file_name")
#view(combined_data)
```




# Social variables

`Social variable` dataset was extracted from the participants [ICE-Nigeria]( http://ice-corpora.net/ice/index.html) to account for how social variables (e.g., `age`, `profession`, `gender`) 

## Step 1

Read the data
```{r load social variable dataset}
nurse_social_var <- read_excel("nurse_social_var.xlsx",
                                     na=c("", "Na", "NA", "N/A", "n/a", "na", "-"))
head(nurse_social_var)
```

## Step 2

**Re-categorize** the `profession` column into `broadcaster` and `politician` rather than the original column that has `Radio presenter`, `TV presenter`, `TV journalist`, and `Politician`

```{r}
nurse_social_var <- nurse_social_var %>%
  filter(Profession %in% c("Radio presenter", "TV presenter", "TV journalist", "Politician"))%>%
  mutate(Profession_category = ifelse(Profession %in% c("Radio presenter", "TV presenter", "TV journalist"), "Broadcaster", "Politician"))
head(nurse_social_var)
```

## Step 3

**Re-categorize** the `age` column into `younger` and `older` rather than the original column that has `numerical values`. The original column has between 20 and 75 years. The new category groups ages between 20 and 45 as younger, 46 and above as older
```{r}
nurse_social_var <- nurse_social_var %>%
  filter(Age %in% c(23, 47, 37, 76, 54, 53, 42, 46))%>%
  mutate(Age_category = ifelse(Age %in% c(23, 37, 42), "younger", "older"))
head(nurse_social_var)
```

# What are the effects of social variables on NURSE vowel?

To calculate the effect of social variables on NURSE vowel, `nurse_social_var` and `nurse_segment_info` data frames are joined by files.
```{r}
#nurse_soc_segment <- merge(nurse_segment_info, nurse_social_var, by = "files")
```


The dataframe **nurse_soc_segment** has segment info such as `vowel`, `duration_ms`, start_dur_sec`, `end_dur_sec`, `word`, `word_interval`, `dur_word_start`, `dur_word_end`, `previous_word`, `next_word`, `gender`, `Age_category`, `Profession_category`. 


# Specific questions to answer in the analysis(Proposed analysis)

A. **Phonetic features**

1. What are the NURSE vowel production (any variation) in NE? (**which** form is the most frequent in NE variety?) (This identifies how NURSE vowels in NE differs from other English varieties (such as British or American English?)
- What information does this study reveal on the production of NURSE vowel that differs from the previous studies?

2. What are the formant trajectories of NURSE vowels in NE? (*What* insight does multiple selection of *formants* across time reveal?)
 
3. Aside the *formant* which other acoustic cues explain NURSE peculiarities in NE?
  - duration ~ 1 + vowel (1|word) 
  - intensity ~ 1 + vowel (1|word)

4. Do the cutoff for winning formants vary by vowel, gender, ethnicity, or profession? This may further reveal variables that trigger complex in NE NURSE vowels?

B. **Context**
Does context determine NURSE vowel production in NE?
    I expect to see something interesting. 
    duration ~ 1 + (1+vowel|word) (1|speaker)
    *does* the context(`phonological environment`) determine phoneme production for NURSE vowel in NE?
    f3 ~ 1 + (1+vowel|word) (1|speaker)
    *To what extent* does `f3` describe NURSE vowel production in different phonological contexts? (is rhoticization a cue for NURSE vowel in NE)
    
C. **Soical variable (Mixed Effects)**
model 1 = (duration ~ 1 + Age_category + Gender + Ethnicity + Profession_category) (1|Word) + (1|file_name)
model 2 = (duration ~ 1 + Age_category + Gender + Ethnicity + Profession_category) (1|Word) + (1|file_name)
model 3 = (f0 ~ 1 + Age_category + Gender + Ethnicity + Profession_category) (1|Word) + (1|file_name))
(aggregate)
model 3(formant) = f3 (previous studies show *f3* as a cue for *categorical* or *gradient* rhoticization) What does *f3* reveal in this study?  
 (f3 ~ 1 + Vowels)



# Model comparison

# tidy_csv files for analysis
```{r}
write_csv(nurse_social_var, "nurse_social_var.csv")
write_csv(nurse_raw_aggregated, "nurse_raw_aggregated.csv")
write_csv(nurse_segment_info, "nurse_segment_info.csv")
write_csv(nurse_vowel, "nurse_vowel.csv")
```

# Session info
```{r}
sessionInfo()
```