---
title: "Data Pipeline"
author: "Oluwasegun Amoniyan"
date: "03/27/2024"
output: 
  github_document: 
    toc: TRUE
---

```{r setup}
##Set knitr options (show both code and output, show output w/o leading #)
knitr::opts_chunk$set(echo = TRUE, include = TRUE, comment=NA)

#load tidyverse
library(tidyverse)

#load excel
library(readxl)
```

# loading raw files
The first file (as nurse_vowel) includes **nurse_vowel** that contains **file_name**, **formant trajectories**, **formant bindwidth**, **pitch**, **intensity** and **harmonity**
The second file (as nurse_segment_info) includes *segment information** 

```{r}
#load raw file
nurse_vowel <- read_excel("nurse_vowel.xlsx",
                                     na=c("", "Na", "NA", "N/A", "n/a", "na", "-", "undefined"))

nurse_raw_aggregated <- read_excel("nurse_raw_aggregated.xlsx",
                                     na=c("", "Na", "NA", "N/A", "n/a", "na", "-", "undefined"))

unstructured_seg_info <- read_csv("unstructured_seg_info.csv",
                                     na=c("", "Na", "NA", "N/A", "n/a", "na", "-", "undefined"))
```


#rename the first column in `nurse_raw_aggregated`before joining the dataframes together.

# Working on `nurse_segment_info`
Data cleaning
  **remove** uninformative columns including `omit`, `previous_sound` and `next_sound`

```{r}
nurse_raw_aggregated <- 
  nurse_raw_aggregated %>%
 rename(
    file_name = contains("þÿfile"))
head(nurse_raw_aggregated)
```

# Tidy unstructed_seg_info

`unstructed_seg_info` has segment features for NURSE vowels.  
```{r}
unstructured_seg_info <- read.csv("unstructured_seg_info.csv")

nurse_segment_info <- separate(unstructured_seg_info, inputfile.outputfile.vowel.interval.duration.start.end.previous_sound.next_sound.omit.word.word_interval.word_start.word_end.previous_word.next_word, 
                into = c("inputfile", "outputfile", "vowel", "interval", "duration", "start", "end", "previous_sound", "next_sound", "omit", "word", "word_interval", "word_start", "word_end", "previous_word", "next_word"), 
                sep = ",")
```
```{r}
#before joining nurse_segment_info with the aggregate by outputfile, there must be a similar name between the two dataframes 
nurse_segment_info <- 
  nurse_segment_info %>%
 rename(
    file_name = contains("outputfile"))
head(nurse_segment_info)
```


**NURSE vowel production by audio files**

```{r}
nurse_segment_info %>%
  select(inputfile)%>%
  table()
```



# remove uninformative columns in `nurse_segment_info`.
The following columns(`previous_sound`, `next_sound` and `omit`) are dropped in the `nurse_segment_info` because they add no value to the `dataframe`.


```{r}
nurse_segment_info <- nurse_segment_info %>%
  select(-c("previous_sound", "next_sound", "omit"))
```

#the data `nurse_segment_info` is now tidy removing uninformative columns 
```{r NURSE vowels in NE}
nurse_segment_info%>%
  select(vowel)%>%
  table()
```

# Combine files
combine both files as a single `csv` file before `data wrangling`

```{r}
#combined_data <- merge(nurse_segment_info, nurse_raw_aggregated, by = "file_name")
#view(combined_data)
```




# Social variable
`Social variable` dataset was extracted from the participants [ICE-Nigeria]( http://ice-corpora.net/ice/index.html) to account for how social variables (e.g., `age`, `profession`, `gender`) 

## Step 1
Read the data
```{r load social variable dataset}
nurse_social_var <- read_excel("nurse_social_var.xlsx",
                                     na=c("", "Na", "NA", "N/A", "n/a", "na", "-"))
head(nurse_social_var)
```

## Step 2
**Re-categorize** the `profession` column into `broadcaster` and `politician` rather than the original column that has `Radio presenter`, `TV presenter`, `TV journalist`, `Politician`

```{r}
nurse_social_var <- nurse_social_var %>%
  filter(Profession %in% c("Radio presenter", "TV presenter", "TV journalist", "Politician"))%>%
  mutate(Profession_category = ifelse(Profession %in% c("Radio presenter", "TV presenter", "TV journalist"), "Broadcaster", "Politician"))
head(nurse_social_var)
```

## Step 3
**Re-categorize** the `age` column into `younger` and `older` rather than the original column that has `numerical values`. The original column has between 20 and 75 years. The new category groups ages between 20 and 45 as younger, 46 and above as older
```{r}
nurse_social_var <- nurse_social_var %>%
  filter(Age %in% c(23, 47, 37, 76, 54, 53, 42, 46))%>%
  mutate(Age_category = ifelse(Age %in% c(23, 37, 42), "younger", "older"))
head(nurse_social_var)
```
# What are the effects of social variables on NURSE vowel?
To calculate the effect of social variables on NURSE vowel, `nurse_social_var` and `nurse_segment_info` data frames are joined by files.
```{r}
#nurse_soc_segment <- merge(nurse_segment_info, nurse_social_var, by = "files")
```


The dataframe **nurse_soc_segment** has segment info such as `vowel`, `duration_ms`, start_dur_sec`, `end_dur_sec`, `word`, `word_interval`, `dur_word_start`, `dur_word_end`, `previous_word`, `next_word`, `gender`, `Age_category`, `Profession_category`. 







# Specific questions to answer in the analysis(Proposed analysis)

A. **Phonetic features**

1. What are the NURSE vowel production (any variation) in NE? (**which** is the over winner in NE variety?) (This identifies how NURSE vowels in NE differs from other English varieties (such as British or American English?)
- What information does this study reveal on the production of NURSE vowel that differs from the previous studies?

2. What are the formant trajectories of NURSE vowels in NE? (*What* insight does multiple selection of *formants* across time reveal?)
 
3. Aside the *formant* which other acoustic cues explain NURSE peculiarities in NE?
  - duration ~ 1 + vowel (1|word) 
  - intensity ~ 1 + vowel (1|word)

4. Do the cutoff for winning formants vary by vowel, gender, ethnicity, or profession? This may further reveal variables that trigger complex in NE NURSE vowels?

B. **Context**
Does context determine NURSE vowel production in NE?
    I expect to see something interesting. 
    duration ~ 1 + (1+vowel|word) (1|speaker)
    *does* the context(`phonological environment`) determine phoneme production for NURSE vowel in NE?
    f3 ~ 1 + (1+vowel|word) (1|speaker)
    *To what extent* does `f3` describe NURSE vowel production in different phonological contexts? (is rhoticization a cue for NURSE vowel in NE)
    
C. **Soical variable (Mixed Effects)**
model 1 = (duration ~ 1 + Age_category + Gender + Ethnicity + Profession_category) (1|Word) + (1|file_name)
model 2 = (duration ~ 1 + Age_category + Gender + Ethnicity + Profession_category) (1|Word) + (1|file_name)
model 3 = (f0 ~ 1 + Age_category + Gender + Ethnicity + Profession_category) (1|Word) + (1|file_name))
(aggregate)
model 3(formant) = f3 (previous studies show *f3* as a cue for *categorical* or *gradient* rhoticization) What does *f3* reveal in this study?  
 (f3 ~ 1 + Vowels)


```{r}
# Read the CSV file
nurse_segment_info <- read.csv("unstructured_seg_info.csv")


nurse_segment_info <- separate(nurse_segment_info, inputfile.outputfile.vowel.interval.duration.start.end.previous_sound.next_sound.omit.word.word_interval.word_start.word_end.previous_word.next_word, 
                into = c("inputfile", "outputfile", "vowel", "interval", "duration", "start", "end", "previous_sound", "next_sound", "omit", "word", "word_interval", "word_start", "word_end", "previous_word", "next_word"), 
                sep = ",")

nurse_segment_info %>%
  select(inputfile)%>%
  table()


# Write the reshaped data to a new CSV file
write.csv(nurse_segment_info, "segment_info.csv")
```
# Model comparison

# tidy_csv files for analysis
```{r}
write_csv(nurse_social_var, "nurse_social_var.csv")
write_csv(nurse_raw_aggregated, "nurse_raw_aggregated.csv")
write_csv(nurse_segment_info, "nurse_segment_info.csv")
```

# Session info
```{r}
sessionInfo()
```