---
title: "Sociophonetic study of NURSE vowels in Nigerian English"
course: LING 2020 - Data Science Research in Linguistics
author: "Oluwasegun Amoniyan"
format: pptx
editor: visual
---

## Motivation for the study

`NURSE vowel` means `vowels that are produced` in similarity or variation to the mid-central vowel `[ɜ]` as `[nɜs]`

The existing literature contains information about \`NURSE vowel\` production, particularly in the English inner circle. However, their findings often overgeneralize other Englishes without providing supporting evidence. This study will investigate the phenomenon in the outer circle of English and provide a balanced description of the variation that characterizes the `NURSE vowel`.

## Background and Review

### Background

**NURSE vowels** have received scholarly attention in different *varieties* (especially in the inner circle) of English (e.g., *American English*, *British English*, *New Zealand English*). Their results are similar.

### Review

Studies of NURSE vowel variation across several English varieties in monolingual or bilingual contexts show merging (as a centralized vowel) or maintaining a three-way contrast realized as \[ɪ\], \[ɛ\], and \[ʌ\], with the varied presence of rhotacization. These realizations are influenced by social (e.g. **gender**, **age**) and linguistic factors (e.g., **speech style**) (Li et. al., 2021; Maclagan et. al., 2017; Mayr, 2010; Mesthrie & Chevalier, 2014; Watson & Clark, 2013).

## Research gap

However, very little is known about NURSE vowel variation in New Englishes like Nigerian English (NE), which exists in a rich multilingual context with regional varieties (i.e., Hausa, Igbo, Yoruba) and over 250 ethnic groups. This study bridges the gap in the knowledge of NURSE vowel production in Englishes.

-   What are NURSE vowel productions in NE?
-   What possible factors determine variation in NURSE vowel production in NE?
-   To what extent do phonetic features describe vowel identity of NURSE vowels?

## Research Hypothesis

The production of NURSE vowels in Nigerian English may vary across **linguistic** (vowel production and phonological context), **ethnicity** (Hausa, Igbo, and Yoruba), and **sociolinguistic** (age, gender, profession) factors. This hypothesis explains the effect of multilingual and diverse socio-cultural spaces on NURSE vowel production in Nigerian English. Specifically:

-   NURSE vowel may have variant production in NE
-   Phonological context may affect NURSE vowel production in NE
-   Nigeria has over 200 ethnic groups and three major (Hausa, Igbo, and Yoruba); I hypothesize that ethnicity may determine NURSE vowel production in NE
-   Sociolinguistic factors may determine how NURSE vowels are produced in NE These factors may contribute to the variation in the production of NURSE vowels than in monolingual speech context.

# Methodology

## Data [ICE-Nigeria](http://ice-corpora.net/ice/index.html)

The data for this study were drawn from the International Corpus of English Nigeria (ICE-Nig). The corpus contains 1,010,382 collections, with 609,586 spoken words and 400,796 written words (Wunder et al., 2010). The sound and text files were downloaded and uploaded to WebMAUS for forced alignment. The forced-aligned speech samples were manually adjusted on the fourth tier on [Praat](https://www.fon.hum.uva.nl/praat/download_mac.html).

-   16 sound files (10 - `bnews`)
-   two speech styles (*broadcast talk* and *broadcast news*)
-   All speakers in this study used NE as second language users [participants](https://sourceforge.net/projects/ice-nigeria/files/latest/download)
-   Vary by gender (`male` vs `female`), `ethnicity`, `speech style`, `Profession`, and `age`

```{r warning=FALSE}
library(tidyverse)
library(lme4)
library(misty)
library(sjPlot)

nurse_segment <- read_csv("C:/Users/oamon/Oluwasegun DS Project/Sociophonetic-study-of-NURSE-vowels-in-NE/tidy_csv_files/nurse_segment_info.csv",
                                     na=c("", "Na", "NA", "N/A", "n/a", "na", "-", "undefined"))

nurse_segment%>%
  select(inputfile)%>%
  table()
```

## Fast Track

[Fast Track](https://github.com/santiagobarreda/FastTrack/tree/master/Fast%20Track) reports 5 formant levels (f1, f2, f3, f4, f5), with the lowest as 4700Hz and the highest as 7550Hz. The algorithm generates 20 formant for undistorted formants and selects a winner after comparison with other possible formant outputs. Unlike previous studies of NURSE vowels that implemented midpoint and three-point vowel measurement, Fast Track would enable me account for formant trajectories. This algorithm has better formant representation than midpoint and three-point vowel measurement. [Samples](https://github.com/ClassOrg-Data-Sci-2024/Sociophonetic-study-of-NURSE-vowels-in-Nigerian-English/blob/main/file_images/images_file_comparisons/bnew_01_images_comparison/bnew_01_0001_comparison.png)

[segment info](https://github.com/ClassOrg-Data-Sci-2024/Sociophonetic-study-of-NURSE-vowels-in-Nigerian-English/blob/main/tidy_csv_files/nurse_segment_info.csv) [nurse aggregate](https://github.com/ClassOrg-Data-Sci-2024/Sociophonetic-study-of-NURSE-vowels-in-Nigerian-English/blob/main/tidy_csv_files/nurse_raw_aggregated.csv) [nurse social](https://github.com/ClassOrg-Data-Sci-2024/Sociophonetic-study-of-NURSE-vowels-in-Nigerian-English/blob/main/tidy_csv_files/nurse_social_var.csv)

### Data Wrangling

-   remove uninformative columns
-   adding and renaming some columns
-   categorization of age group, profession
-   joining data frames

## Vowel normalization

***Bark difference metrics***

-   to eliminate physiological differences among speakers
-   to preserve socio-linguistic/dialectal/cross-linguistic differences in vowel quality
-   to preserve phonological distinctions among vowels [navigate](http://lingtools.uoregon.edu/norm/norm1_methods.php)

## Statistical Analysis

The [raw files](https://github.com/ClassOrg-Data-Sci-2024/Sociophonetic-study-of-NURSE-vowels-in-Nigerian-English/tree/main/raw_Data) were analyzed using Mixed effects regression with **tidyverse** and **lme4** to account for linguistic, ethnic and social variables that determine (or may contribute to) variation in Nigerian English NURSE vowels.

# Analysis

## NURSE vowel productions in NE

```{r}
#NURSE_vowel <- nurse_segment%>%
 # select(vowel)%>%
  #count()%>%
  #print

#NURSE_vowel <- nurse_segment%>%
 # select(vowel)%>%
  #table()/246*100
#freq/overall sum*100 to estimate the percentages
#head(NURSE_vowel)

nurse_segment <- nurse_segment %>% 
  mutate(vowel_variation = case_when(
    vowel %in% c("ɒ", "ɔ") ~ "mid_back",
    vowel %in% c("æ", "ɑ") ~ "low_front_back",
    vowel %in% "ɛ" ~ "mid_front",
    vowel %in% "ɪ" ~ "high_front",
    vowel %in% "ɜ" ~ "mid-central",
    TRUE ~ "Other"
  ))

#vowel_variation <- nurse_segment%>%
 # select(vowel_variation)%>%
  #table()/246*100
#head(vowel_variation)


ggplot(nurse_segment, aes(x = vowel_variation, fill = vowel)) +
  geom_bar() +
  labs(x = "vowel", y = "Frequency", color = "Vowel") +
  ggtitle("Frequencies of NE vowel variation (with)out length contrast in NE") +
  theme_minimal()
```

## Duration patterns for NURSE vowels in NE

This tests if there is a merger by duration (not by vowel space as largely reported in the literature)

```{r}
NURSE_vowel_dur <- nurse_segment%>%
  group_by(vowel)%>%
  summarize(mean_dur_vowel = mean(duration))
#head(NURSE_vowel_dur)

ggplot(nurse_segment, aes(x=vowel, y = duration)) +
         geom_point() +
  labs(x = "vowel", y = "duration") +
  ggtitle("Duration patterns of NE vowel in NE") +
  theme_minimal()
```

The duration distinguishes tense from lax vowels. The average duration of vowel production in a NURSE context in Nigerian English is 13ms (0.126) for `[æ]` and 14ms (0.14) for `[ɑ]`. Similarly, the mean duration of the vowels `[ɒ]`, `[ɔ]`, `[ɛ]`, and \[ɜ\] range from 11ms (0.113) to 13ms (0.130). [Segment info](https://github.com/ClassOrg-Data-Sci-2024/Sociophonetic-study-of-NURSE-vowels-in-Nigerian-English/blob/main/tidy_csv_files/segment_info.csv) (Is duration significant?)

#### Regression analysis of duration

Duration **does not** significantly determine variance in NURSE vowel. The model, the results show that the realization of the NURSE vowel as a low back vowel (`[ɑ]`) was statistically significantly different in `duration` from other NURSE vowels (intercept = 0.13, 95%, t(237) = 12.54, p \< .001). Meanwhile, the average duration for `[æ]` was 8ms longer than \[a:\], though the relationship was not statistically significant (p\>0.05). This may indicate that duration is unreliable for differentiating NURSE vowels in NE. Rather, vowel height may help to differentiate NURSE vowel phoneme.

```{r}
#regression model for duration
duration_model <- lmer(duration ~ 1 + vowel + (1|word), data = nurse_segment)
#duration_model %>% summary()

#let's the coeffiecient of NURSE vowels across words
NURSE_context_dur <- coef(duration_model)%>%
  pluck('word')%>%
  rownames_to_column('NURSE_tokens')

#load the coefplot package
library(coefplot)

# Create the coefficient plot
coefplot(duration_model)

```

## Phonological effect on NURSE vowel

One of the [hypotheses](https://github.com/ClassOrg-Data-Sci-2024/Sociophonetic-study-of-NURSE-vowels-in-Nigerian-English/blob/main/research%20hypothesis.md) for the study identifies the effect of phonological environment on NURSE vowel production in NE. To attempt this task, we need to create a column `nurse_determinant` where we assign numeric to each level of the `nurse_vowel_space` for `Poisson analysis`. The column has three levels: NURSE vowels realized as `front_vowel`, `back_vowel,` and `central_vowel`. The numeric \[`front_vowel = 0`; `back_vowel = 1`, `central_vowel = 2`\]. This model will reveal whether there is any relationship between `vowel space` and `words` in NE.

```{r}
nurse_segment <- nurse_segment %>% 
  mutate(nurse_vowel_space = case_when(
    vowel %in% c("ɒ", "ɔ", "ɑ") ~ "back_vowel",
    vowel %in% c("æ", "ɛ", "ɪ") ~ "front_vowel",
    vowel %in% "ɜ" ~ "central_vowel",
    TRUE ~ "Other"
  ))

#nurse_segment%>%
 # select(nurse_vowel_space)%>%
  #table()/246*100
#phonological environment effect
nurse_segment <- nurse_segment %>% 
  mutate(nurse_determinant = case_when(
    nurse_vowel_space %in% "front_vowel" ~ "0",
    nurse_vowel_space %in% "back_vowel" ~ "1",
    nurse_vowel_space %in% "central_vowel" ~ "2",
      TRUE ~ "Other"
  ))

#make the categorical column numeric to satisfy poisson assumption
nurse_segment <- nurse_segment %>%
  mutate(nurse_determinant = as.numeric(nurse_determinant))

#regression analysis for the `phonological effect` on NURSE vowel production
model_nurse_determinant1 <- glmer(nurse_determinant ~ 1 + (1|word), data = nurse_segment, family = poisson)
#summary(model_nurse_determinant1)

ggplot(nurse_segment, aes(x = vowel, fill = word)) +
  geom_bar() +
  labs(x = "NURSE vowels", y = "Freq of NURSE vowels")+
   ggtitle("Frequencies of NURSE vowel production by word")
```

\`Phonological environment\` determines NURSE vowel production in NE. The random effect for \`word\` indicates that the influence of \`word\` on vowel production varies. The SD (0.28) estimates the degree of variability among words. Furthermore, the fixed effect estimate for the intercept (-0.58) indicates the NURSE vowel productions when no word-specific factor is present. This support the evidence that the production of vowels in NE is determined by the \`phonological environment\` that NE speakers utter, with various words having variable impacts on vowel identity. This is novel and may not be general across Englishes.

## NURSE vowel formant trajectories

```{r}
##Formant trajectory
nurse_aggregate <- read_csv("tidy_csv_files/nurse_raw_aggregated.csv",
                                     na=c("", "Na", "NA", "N/A", "n/a", "na", "-", "undefined"))
#head(nurse_aggregate)

#Another data frame
nurse_segment <- 
  nurse_segment %>%
 rename(
    file_name = contains("outputfile"),
    files = contains("inputfile"))
#head(nurse_segment)

#combined the dataframe by `file_name`
segment_aggregate <- merge(nurse_aggregate, nurse_segment %>% select(c("vowel", "word", "file_name", by = "file_name")))

#remove columns that are not needed
segment_aggregate <- segment_aggregate %>%
  select(-c("files", "label", "group", "number", "f0"))

#rename `file_name` as `id` as a concise name for vowel grouping
segment_aggregate <- 
  segment_aggregate %>%
 rename(
    id = contains("file_name"))


my_nurse_vowel_formant <- segment_aggregate %>%
    # Take the median of each vowel at each timepoint
  summarize(across(starts_with("f"), median, na.rm = TRUE), .by = vowel) %>%
  
  # Convert it into a "very tall" format
  pivot_longer(cols = starts_with("f"), names_to = "formant_percent", values_to = "hz") %>%
 separate_wider_position(formant_percent, 
                          widths=c(formant = 2, timepoint = 1))%>%
  
  # unite the columns and view
  unite(id, formant, vowel, remove = FALSE)


# Define the formula as a function
bark_norm <- function(fi) {
  zi <- 26.81 / (1 + 1960 / fi) - 0.53
  return(zi)
}

my_nurse_vowel_formant <- my_nurse_vowel_formant %>%
  mutate(formant_bark_norms = sapply(hz, bark_norm))
#head(my_nurse_vowel_formant)

#visualization
ggplot(my_nurse_vowel_formant, aes(x = timepoint, y = formant_bark_norms, group = id, color = formant)) +
  geom_line() + 
  facet_wrap(~vowel, nrow=3)+
  ggtitle("Formant trajectories for NURSE vowels in NE")+
  theme_classic()
```

### NURSE vowel production in NE by age and gender

```{r}
nurse_social <- read_csv("tidy_csv_files/nurse_social_var.csv",
                                     na=c("", "Na", "NA", "N/A", "n/a", "na", "-", "undefined"))
#nurse_social dataframe has only social variables and need to join `nurse_segment` to account for the effect of social variables on NURSE vowels.

#rename `files` in `nurse_social` dataframe as `id` in order to join both dataframe by `id`. 

nurse_social <- nurse_social %>%
  rename(
    id = contains("files"))
#new dataframe as `nurse_social_segment` has both `nurse_social` and `nurse_segment` merged by `id`
nurse_segment <- 
  nurse_segment %>%
  rename(
    id = contains("files"))

nurse_social_segment <- merge(nurse_segment, nurse_social, by = "id")
#head(nurse_social)
#to view
#head(nurse_social_segment)

#frequency count of the NURSE vowel production by age
nurse_freq_age <- table(nurse_social_segment$vowel, nurse_social_segment$Age_category)
#head(nurse_freq_age)
#percentage
nurse_age_vowel <- nurse_social_segment %>%
  select(Age_category, Gender, vowel)%>%
  table()/246*100
#head(nurse_age_vowel)

ggplot(nurse_social_segment, aes(x = vowel, fill = Gender)) +
  geom_bar() +
  facet_wrap(vars(Age_category), ncol = 3) +
  labs(x = "NURSE vowels", y = "Freq of NURSE vowels")+
   ggtitle("Frequencies of NURSE vowel production by social group")
```

### NURSE vowel determinant by age and gender

```{r}
ggplot(nurse_social_segment, aes(x = vowel, fill = word)) +
  geom_bar() +
  facet_wrap(vars(Age_category), ncol = 3) +
  labs(x = "NURSE vowels", y = "Freq of NURSE vowels")+
   ggtitle("Frequencies of NURSE vowel production by age_category")
```

Age(sign), gender(not sig, no conv)

```{r}
model_nurse_age_det <- glmer(nurse_determinant ~ 1 + Age_category + (1|word), data = nurse_social_segment, family = poisson)
#summary(model_nurse_age_det)
#model_nurse_age_det %>% tab_model()

plot_model(model_nurse_age_det, type = "pred", terms = c("Age_category"))
```

### Duration patterns of NURSE vowel production in NE by age and gender

Sign by `gender` (alone)

```{r}
#nurse_segment <- nurse_segment %>%
 # mutate(duration.cen = center(duration, value=min(duration)))

#nurse_social_segment %>% group_by(Age_category) %>% summarize(age_vdur = mean(duration.cen))

age_nurse_vowel <- lmer(duration ~ 1 + Age_category + Gender + (1|vowel), nurse_social_segment) #age_nurse_vowel %>% summary() age_nurse_vowel %>% tab_model()

plot(age_nurse_vowel, Gender ~ resid(.), abline = 0 ) 
plot(age_nurse_vowel, Age_category ~ resid(.), abline = 0 ) 
plot(age_nurse_vowel, vowel ~ resid(.), abline = 0 ) 
plot_model(age_nurse_vowel, type = "pred", terms = c("Age_category", "Gender", "vowel"))
```

## NURSE vowel production by `ethnicity`

```{r}
ggplot(nurse_social_segment, aes(x = vowel, y = duration), fill = Gender) +
  geom_point() +
  facet_wrap(vars(Ethnicity), ncol = 3) +
  labs(x = "NURSE vowels", y = "Duration of NURSE vowels")+
   ggtitle("Duration patterns NURSE vowel production by social group")
```

```{r}
age_nurse_ethnicity <- lmer(duration ~ 1 + Ethnicity + vowel + (1|word),  nurse_social_segment) 
#age_nurse_ethnicity %>% summary()
#sjPlot:: tab_model(age_nurse_ethnicity)

plot_model(age_nurse_ethnicity, type = "pred", terms = c("Ethnicity", "vowel"))
```

### NURSE vowel determinant by `ethnicity`

```{r}
nurse_ethn_det <- glmer(nurse_determinant ~ 1 + Ethnicity + (1|word), data = nurse_social_segment, family = poisson)
#summary(nurse_ethn_det)
plot_model(nurse_ethn_det, type = "pred", terms = c("Ethnicity"))
plot_model(nurse_ethn_det, type = "pred", terms = c("Ethnicity", "word"))
```

## Discussion and future work

Evidently, there are variations in NURSE vowels other than categorical or gradient r-presence. If we included other speech styles later, the variation might exceed what we observed.

My findings are similar to those of previous studies, which found that `gender`, `age`, `style`/`register`/`community of practice`, and `geographical location` could affect `NURSE vowel production`. (Li et. al., 2021; Maclagan et. al., 2017; Mayr, 2010; Mesthrie & Chevalier, 2014; Watson & Clark, 2013)

My analysis is limited (just 246 data points), so I will add more speech samples before generalization.

As my interest in sociophonetics grows, I will continue to learn R packages and skills in data science for data management, presentation, and accessibility.

## References

Li, Z., Gut, U., & Schützler, O. (2021). nurse Vowels in Scottish Standard English: Still Distinct or Merged? Journal of English Linguistics, 49(3), 305-330.   

Maclagan, M., Watson, C. I., Harlow, R., King, J., & Keegan, P. (2017). Investigating the sound	change in the New Zealand English nurse vowel/ᴈ. Australian Journal of Linguistics, 37(4), 465-485.   

Mayr, R. (2010). What exactly is a front-rounded vowel? An acoustic and articulatory 

investigation of the NURSE vowel in South Wales English. Journal of the International Phonetic Association, 40(1), 93-112.   

Mesthrie, R., & Chevalier, A. (2014). Sociophonetics and the Indian diaspora: The NURSE vowel and other selected features in South African Indian English. In English in the Indian diaspora (pp. 85-104). John Benjamins.   
