
---
title: "NURSE Analysis"
author: "Oluwasegun Amoniyan"
date: "04/06/2024"
output: 
  github_document: 
    toc: TRUE
---

```{r setup}
##Set knitr options (show both code and output, show output w/o leading #)
knitr::opts_chunk$set(echo = TRUE, include = TRUE, comment=NA)

#load packages
library(tidyverse)
library(ggplot2)
#library(reshape2)
library(misty)
library(lme4)
library(lmerTest) 
library(performance)
```

# Analysis of NURSE vowels in Nigerian English (NE)

```{r}
#load `csv` that has segment information
nurse_segment <- read_csv("C:/Users/oamon/Oluwasegun DS Project/Sociophonetic-study-of-NURSE-vowels-in-NE/Analysis/tidy_csv_files/nurse_segment_info.csv",
                                     na=c("", "Na", "NA", "N/A", "n/a", "na", "-", "undefined"))

#rename `outputfile` and `inputfile` as `file_name` and `id` respectively
#nurse_segment <- 
 # nurse_segment %>%
 #rename(
  #  file_name = contains("outputfile"),
   # id = contains("inputfile"))
```
## NURSE vowel production in NE

### Variation in NURSE vowel 

1a. What are the distinct NURSE vowel production (any variation) in NE? (**which** variant is the most frequent in NE variety and does the variance describe how NURSE vowels in NE differs from other English varieties (such as British or American English?)

Here, we identify variant phonemes that are produced as NURSE vowel in NE. This variation identifies the difference in phoneme by length to get specific NURSE vowel production.

```{r NURSE vowels in NE}
#number of NURSE vowels extracted
NURSE_vowel <- nurse_segment%>%
  select(vowel)%>%
  count()%>%
  print

NURSE_vowel <- nurse_segment%>%
  select(vowel)%>%
  table()/246*100
#freq/overall sum*100 to estimate the percentages
head(NURSE_vowel)
```
The total number of NURSE vowels extracted 246 tokens (too small to generalize). Describution: 

`[ɜ]` is realized as `[nɜs]` by 6.4%, as `[na:s]` by 13%, as `[nas]` by 27%, as `[nɔs]` by 17%,  as `[nɔ:s]` by 16%, as `[nɛs]` by 20%, and as `[ɪ]` by 0.4%. The overall results show that the variants with shorter duration are more favored than the longer ones in NE variety. Each of these productions has segment info [here](https://github.com/ClassOrg-Data-Sci-2024/Sociophonetic-study-of-NURSE-vowels-in-Nigerian-English/blob/main/tidy_csv_files/segment_info.csv) and each segment has predicted and winning formants [here](https://github.com/ClassOrg-Data-Sci-2024/Sociophonetic-study-of-NURSE-vowels-in-Nigerian-English/tree/main/file_images/images_file_comparisons).  

### NURSE vowel variation without length contrast

Here, we identify variant phonemes that are produced as NURSE vowel in the variety. This variation does not first consider the difference in phoneme by length to get the general overview.

Before analyzing the variation in NURSE vowel in NE, let's create a column (`vowel_variation`) that does not differentiate NURSE vowels by length.

```{r}
#create a column that has the frequency of the variant form
nurse_segment <- nurse_segment %>% 
  mutate(vowel_variation = case_when(
    vowel %in% c("ɒ", "ɔ") ~ "mid_back",
    vowel %in% c("æ", "ɑ") ~ "low_front_back",
    vowel %in% "ɛ" ~ "mid_front",
    vowel %in% "ɪ" ~ "high_front",
    vowel %in% "ɜ" ~ "mid-central",
    TRUE ~ "Other"
  ))

vowel_variation <- nurse_segment%>%
  select(vowel_variation)%>%
  table()/246*100
head(vowel_variation)


ggplot(nurse_segment, aes(x = vowel_variation)) +
  geom_bar() +
  labs(x = "vowel", y = "Frequency") +
  ggtitle("NURSE vowel distinction in NE") +
  theme_minimal()
```
Unlike Li et al. (2021) who discovered three-way distinction in Scottish NURSE vowel, the findings revealed that there were five(5) NURSE vowel distinction in NE. The frequency distribution revealed open vowel [e.g., a] as the most frequent variant. Also, this analysis supported Collins & Mees (1990) who asserted that NURSE vowel could vary by vowel space. The distribution of Figure 1 revealed low front and back as the most frequent variant of NURSE, followed by mid-back and mid-front. This result could inform us that NE speakers often produced NURSE vowels as low (front/back), mid_back or mid_front vowels with different frequency.  

As Penhallurick (2008) suggests that duration plays a role in determining the degree of mouth openness and lip-rounding. 

The results show that `[ɜ]` in NURSE for BrE, AmE, CanE and New Zealand English is often realization as `[a]` with 99 (40%), `[ɔ]` with 81 (32%), `[ɜ]` with 16 (6.5%), and `[ɛ]` with 49 (19%). This shows that similar variant with the inner circle English is 6.5% possible in NE variety; however, it's more produced as [nas], [nɔs] and [nɛs]. The frequent production of [nɜs] as [nas], [nɔs] and [nɛs] in NE reveals that the variation of NURSE vowel is beyond presence or absence of rhotics. If we shift our focus rhotics of NURSE vowel in NE, we will eventually be discussing variant form that is 10% present in NE. 

### NURSE vowel variation by `backness`

Here, we classified NURSE vowel variant that are produced as NURSE vowel in NE as **front**, **central**, and **backness**. This classification does not consider length contrast but identify.

Before analyzing the variation in NURSE vowel in NE, let's create a column (`vowel_variation`) that does not differentiate NURSE vowels by length.
```{r}
nurse_segment <- nurse_segment %>% 
  mutate(nurse_vowel_space = case_when(
    vowel %in% c("ɒ", "ɔ", "ɑ") ~ "back_vowel",
    vowel %in% c("æ", "ɛ", "ɪ") ~ "front_vowel",
    vowel %in% "ɜ" ~ "central_vowel",
    TRUE ~ "Other"
  ))

nurse_segment%>%
  select(nurse_vowel_space)%>%
  table()/246*100

ggplot(nurse_segment, aes(x = nurse_vowel_space)) +
  geom_bar() +
  labs(x = "vowel", y = "Frequency") +
  ggtitle("NURSE vowel distinction NE by vowel position") +
  theme_minimal()
```
The results show the distribution of vowel production in Nigerian English across front, central and back vowel space. Back vowels are frequent, accounting for 45.93% of overall vowel production, whereas front vowels were 47.56%, while the central vowel were the least common,  6.50% in NE. This distribution implies that Nigerian English has a somewhat equal proportion of back and front vowels, with back vowels being less common.

## Duration patterns for NURSE vowels in NE  
```{r}
#check the duration patterns for NURSE vowels in the data
NURSE_vowel_dur <- nurse_segment%>%
  group_by(vowel)%>%
  summarize(mean_dur_vowel = mean(duration))

NURSE_vowel_var <- nurse_segment%>%
  group_by(vowel_variation)%>%
  summarize(mean_vowel_var = mean(duration))

head(NURSE_vowel_dur)
head(NURSE_vowel_var)
```
### NURSE vowel duration by length

The duration distinguishes tense from lax vowels. The average duration of vowel production in a NURSE context in Nigerian English is 13ms (0.126) for `[æ]` and 14ms (0.14) for `[ɑ]`. Similarly, the mean duration of the vowels [ɒ]`, `[ɔ]`, `[ɛ]`, and [ɜ] range from 11ms (0.113) to 13ms (0.130). [Segment info](https://github.com/ClassOrg-Data-Sci-2024/Sociophonetic-study-of-NURSE-vowels-in-Nigerian-English/blob/main/tidy_csv_files/segment_info.csv) (Is duration therefore significant?)


### NURSE vowel duration by vowel space

Each category of `vowel space` represents a unique articulatory position within the vowel space. For example, the **high_front** category has a relatively modest mean vowel duration of 0.04 (4ms), indicating a more limited articulatory range for high front vowels in Nigerian English. In contrast, categories such as `low_front_back`, `mid-central`, `mid_back`, and mid_front` have higher average duration ranging from 0.118 (12ms) to 0.131 (13ms), implying a wider range of articulatory positions for these vowels. Overall, the variability of vowel production in NURSE production in Nigerian English across various vowel categories, shedding light on the phonetic characteristics of this linguistic context and indicating variations not only in vowels but also in duration.

### Regression analysis of duration

Duration **does not seem** to significantly determine the variance in NURSE vowel production in NE. The model, the results show that the realization of NURSE vowel as low back vowel (`[ɑ]`) was statistically significantly differ in `duration` from other other NURSE vowels (intercept = 0.13, 95%, t(237) = 12.54, p < .001). Meanwhile, the average duration for `[æ]` was 8ms longer than [a:], though the relationship was not statistically significant (p>0.05). This may insights that duration is not reliable to differentiate vowels for NURSE production in NE. Rather, vowel height may help to differentiate NURSE vowel phoneme.

```{r}
#normalization duration around the mean before checking for the effect of duration
nurse_segment <- nurse_segment %>%
  mutate(duration.cen = center(duration, value=min(duration)))

#regression model for duration
duration_model <- lmer(duration.cen ~ 1 + vowel + (1|word), data = nurse_segment)
duration_model %>% summary()
library(sjPlot)
duration_model %>% tab_model()
duration_model %>% plot_model()

NURSE_context_dur <- coef(duration_model)%>%
  pluck('word')%>%
  rownames_to_column('NURSE_tokens')
  

#load the coefplot package
library(coefplot)

# Create the coefficient plot
coefplot(duration_model)

#alternative plotting
duration_m <- data.frame(
  vowel = c("[ɒ]", "[æ]", "[ɔ]", "[ɛ]", "[ɜ]", "[ɪ]"),
  Estimates = c(-0.01, 0.00, -0.02, -0.00, -0.01, -0.09),
  CI_Lower = c(-0.04, -0.02, -0.05, -0.02, -0.04, -0.20),
  CI_Upper = c(0.02, 0.02, 0.01, 0.02, 0.02, 0.02),
  p_value = c(0.372, 0.938, 0.161, 0.949, 0.603, 0.119)
)

ggplot(duration_m, aes(x = vowel, y = Estimates, ymin = CI_Lower, ymax = CI_Upper)) +
  geom_point(size = 3, color = "blue") +
  geom_errorbar(width = 0.2, color = "blue") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Effect of Vowels on Duration (cen)",
       x = "Vowel",
       y = "Estimates") +
  theme_minimal()
```
Similarly, I fitted the second level with `word` (as level 2) and model did not reveal any significant relationship therefore I dropped it from the model. The residual revealed variation across `word` and `vowel` production as 4ms (for `word`) and 3ms (for `vowel`). It means that duration of NURSE vowel in the variety is more variant across `word` than `NURSE vowel`


## Effect of phonological environment on the production of NURSE vowels in NE

One of the [hypotheses](https://github.com/ClassOrg-Data-Sci-2024/Sociophonetic-study-of-NURSE-vowels-in-Nigerian-English/blob/main/research%20hypothesis.md) for the study identifies the effect of phonological environment on NURSE vowel production in NE. To attempt this task, we need to create a column `nurse_determinant` where we assign numeric to each level of the `nurse_vowel_space` for `poisson analysis`. The column has three levels: NURSE vowels realized as `front_vowel`, `back_vowel` and `central_vowel`. The numeric [`front_vowel = 0`; `back_vowel = 1`, `central_vowel = 2`]. This model will reveal whether there is any relationship between `vowel space` and `words` in NE.  

```{r}
#Do `words` (and or `speech`) determine NURSE vowel in NE?
nurse_segment <- nurse_segment %>% 
  mutate(nurse_determinant = case_when(
    nurse_vowel_space %in% "front_vowel" ~ "0",
    nurse_vowel_space %in% "back_vowel" ~ "1",
    nurse_vowel_space %in% "central_vowel" ~ "2",
      TRUE ~ "Other"
  ))
```

### Account for the phonological effect in NE

Create a new column to account account for the phonological effect in NE NURSE 
```{r}
# Define a function to check if a word ends with any of the specified endings
check_segment <- function(word, endings) {
  for (ending in endings) {
    if (endsWith(word, ending)) {
      return(TRUE)
    }
  }
  return(FALSE)
}

# List of consonant sounds
consonant_letter <- c("w", "r", "j", "k", "s", "p", "n", "f", "t", "v", "th", "b", "m", "l", "g", "d", "ch", "sh", "c")

# List of vowels
vowels <- c("a", "e", "i", "o", "u")

# List of endings
endings <- c("_er_", "_ur_", "_ir_", "_or_", "_ear_", "_our_",
             "_l", "_g", "_ur", "_ir")

# Create a new column for phonological environment
nurse_segment <- nurse_segment

# Loop through each word in the 'word' column
for (word in nurse_segment$word) {
  # Check if the word ends with a consonant sound
  if (substring(word, nchar(word)) %in% consonant_letter) {
    # Check if the word ends with any of the specified endings
    if (check_segment(word, endings)) {
      nurse_segment$phonological_environment[nurse_segment$word == word] <- "After consonant"
    } else {
      nurse_segment$phonological_environment[nurse_segment$word == word] <- "After consonant (no specified endings)"
    }
  } else {
    nurse_segment$phonological_environment[nurse_segment$word == word] <- "Not after consonant"
  }
}

# Print the updated dataframe
head(nurse_segment)
```

### Regression analysis for nurse determinant by phonological environment
```{r}
#make the categorical column numeric to satisfy poisson assumption
nurse_segment <- nurse_segment %>%
  mutate(nurse_determinant = as.numeric(nurse_determinant))

#regression analysis for the `phonological effect` on NURSE vowel production
model_nurse_determinant <- glmer(nurse_determinant ~ 1 + phonological_environment + (1|inputfile) + (1|word), data = nurse_segment, family = poisson)
summary(model_nurse_determinant)

library(sjPlot)
#Estimate table that explains the independent and the dependent variables.
#formant_model %>% tab_model()
model_nurse_determinant %>% tab_model()
plot_model(model_nurse_determinant, type = "pred", terms = c("phonological_environment"))
```



### Visualization of phological effect in NURSE vowel production
```{r}
ggplot(nurse_segment, aes(x = vowel, fill = phonological_environment)) +
  geom_bar() +
  labs(x = "NURSE vowels", y = "Freq of NURSE vowels")+
   ggtitle("Effect of phonological environment on NURSE vowel by social group")
```

The model determines if `NURSE vowels` are affected by `phonological environment` and the intercept returned significant. The random effect for `word` indicates that the influence of `word` on vowel production varies. The standard deviation (`SD`) for the random effect (0.2752) estimates the degree of variability among words. Furthermore, the fixed effect estimate for the intercept (-0.5792) indicates the log of vowel productions when no word-specific factor is present. This support the evidence that the production of vowels in Nigerian English is determined by the `phonological environment` that NE speakers utter, with various words having variable impacts on vowel formation.

The model was extended to investigate whether vowel production in Nigerian English is influenced by the specific `word` and the `speaker's speech style` [`broadcast interview` vs `broadcast talk`]. The model adds speaker identification `(id)`, indicating that the effects on vowel production vary between words and speakers. The model did not converge and I dropped `id` from the model.

## What information does the variation in NE NURSE vowel explain?

First, the production of NURSE vowel varies in Nigerian English. Unlike previous studies [link](https://www.cambridge.org/core/journals/journal-of-the-international-phonetic-association/article/what-exactly-is-a-front-rounded-vowel-an-acoustic-and-articulatory-investigation-of-the-nurse-vowel-in-south-wales-english/96F7CC3AC905D8F3DED8E011394310A3), [also](https://journals.sagepub.com/doi/full/10.1177/00754242211025586) in `monolingual context` and these studies explain merger/distinction or varied rhotics; this study, therefore, shows that the more diverse a speech community (e.g., `bilingual`, `multilingual`) is, the more variation in **NURSE vowel** production. Previous studies on NURSE vowels discussed categorical or gradient rhotacization and their participants were speakers of English as L1. The production of NURSE vowel in NE (a multi-lingual speech community) has revealed inherent variation in the production of NURSE vowel. To this effect, researchers on inner circle English may have to restrict their generalization that NURSE vowels have categorical (presence or absence of [r]) or gradient rhoticity. The multi-lingual speech communities have variant forms to mean the same phoneme (as NURSE vowel) without any pragmatic implicature. 

## Formant description of NURSE vowels in NE

- What are the formant features of NURSE vowels in NE? 
- What are the formant trajectories of NURSE vowels in NE?
- What insight does multiple selection of formants across time reveal?

```{r}
#`nurse_aggregate.csv` has phonetic features of NURSE vowel production 
nurse_aggregate <- read_csv("C:/Users/oamon/Oluwasegun DS Project/Sociophonetic-study-of-NURSE-vowels-in-NE/Analysis/tidy_csv_files/nurse_raw_aggregated.csv",
                                     na=c("", "Na", "NA", "N/A", "n/a", "na", "-", "undefined"))

#head(nurse_aggregate)

#Another data frame
nurse_segment <- 
  nurse_segment %>%
 rename(
    file_name = contains("outputfile"),
    files = contains("inputfile"))
head(nurse_segment)

#combined the dataframe by `file_name`
segment_aggregate <- merge(nurse_aggregate, nurse_segment %>% select(c("vowel", "word", "file_name", by = "file_name")))

#remove columns that are not needed
segment_aggregate <- segment_aggregate %>%
  select(-c("files", "label", "group", "number", "f0"))

#rename `file_name` as `id` as a concise name for vowel grouping
segment_aggregate <- 
  segment_aggregate %>%
 rename(
    id = contains("file_name"))
```

Vowel normalization and visualization
```{r}
#store the new data to `my_nurse_vowel_formant` 
my_nurse_vowel_formant <- segment_aggregate %>%
    # Take the median of each vowel at each timepoint
  summarize(across(starts_with("f"), median, na.rm = TRUE), .by = vowel) %>%
  
  # Convert it into a "very tall" format
  pivot_longer(cols = starts_with("f"), names_to = "formant_percent", values_to = "hz") %>%
 separate_wider_position(formant_percent, 
                          widths=c(formant = 2, timepoint = 1))%>%
  
  # unite the columns and view
  unite(id, formant, vowel, remove = FALSE)


# Define the formula as a function
bark_norm <- function(fi) {
  zi <- 26.81 / (1 + 1960 / fi) - 0.53
  return(zi)
}

my_nurse_vowel_formant <- my_nurse_vowel_formant %>%
  mutate(formant_bark_norms = sapply(hz, bark_norm))
head(my_nurse_vowel_formant)

#visualization
ggplot(my_nurse_vowel_formant, aes(x = timepoint, y = formant_bark_norms, group = id, color = formant)) +
  geom_line() + 
  facet_wrap(~vowel, nrow=2)+
  ggtitle("Formant trajectories for NURSE vowels in NE")+
  theme_classic()
```




### Regression analysis of formant features and trajectories

```{r}
#formant
formant_model <- lmer(formant_bark_norms ~ 1 + timepoint + (1|formant), data = my_nurse_vowel_formant)
formant_model %>% summary()
library(sjPlot)
#Estimate table that explains the independent and the dependent variables.
formant_model %>% tab_model()
formant_model %>% plot_model()
#plot_model(formant_model, type = "pred", terms = c("timepoint", "formant"))
```
This analysis shows the estimates, confidence intervals (CI), and p-values for `formant` with normalization. The intercept `timepoint 1` is estimated to be -0.97 with a 95% confidence interval between -0.99 and -0.96, and the p-value is <0.001, indicating high statistical significance. The other estimates (f2, f3, and f4) are 0.82, 1.37, and 1.70, respectively, with very narrow confidence intervals and highly significant p-values (<0.001). This indicates a strong positive association between each formant and the formant levels. 

The random effects analysis shows minimal variance both within and between time points, with a variance estimate `σ^2` of 0.01 and a variance estimate for the intercept across time points (τ00) of 0.01. The intra-class correlation coefficient (ICC) is estimated to be 0.38, indicating that time-point differences account for a moderate proportion of total variance. The model fit is excellent, with a marginal R^2 `0.998` and a conditional R^2 `0.999`. This indicates that the predictors (i.e., formant) explain a significant proportion of the variance in the outcome variable (`formant levels`), both independently and in conjunction with other predictors.



## Do social factors determine production of NURSE vowels in NE?

Apart from the phonetic properties [as stated in the hypothesis](https://github.com/ClassOrg-Data-Sci-2024/Sociophonetic-study-of-NURSE-vowels-in-Nigerian-English/blob/main/research%20hypothesis.md) that described variation in Nigerian English NURSE vowels, we expect social factors such as `age`, `gender` and `profession` to trigger variation in NURSE vowel production.

```{r}
#load csv with social variables
nurse_social <- read_csv("C:/Users/oamon/Oluwasegun DS Project/Sociophonetic-study-of-NURSE-vowels-in-NE/Analysis/tidy_csv_files/nurse_social_var.csv",
                                     na=c("", "Na", "NA", "N/A", "n/a", "na", "-", "undefined"))
#nurse_social dataframe has only social variables and need to join `nurse_segment` to account for the effect of social variables on NURSE vowels.

#rename `files` in `nurse_social` dataframe as `id` in order to join both dataframe by `id`. 

nurse_social <- nurse_social %>%
  rename(
    id = contains("files"))


#rename `outputfile` and `inputfile` as `file_name` and `id` respectively
nurse_segment <- 
  nurse_segment %>%
 rename(
    file_name = contains("outputfile"),
    id = contains("files"))

#new dataframe as `nurse_social_segment` has both `nurse_social` and `nurse_segment` merged by `id`
nurse_social_segment <- merge(nurse_segment, nurse_social, by = "id")
#to view
head(nurse_social_segment)
```

### NURSE vowel production in NE by age and gender

The table (r code below) showed the production of NURSE vowels by age groups (`older` and `younger`). 17 `older` participants produced the vowel low back vowel while only 6 younger participants produced the same vowel. 8 older and 10 younger participants produced NURSE vowel as [ɒ]. Notably, older participants produced the vowel "æ" more frequently than younger participants, with 41 occurrences versus only three. These findings shed light on differences in the production of NURSE vowels across age groups, indicating that age may influence the production of NURSE vowel in Nigerian English.

```{r}
#frequency count of the NURSE vowel production by age
nurse_freq_age <- table(nurse_social_segment$vowel, nurse_social_segment$Age_category)
head(nurse_freq_age)
#percentage
nurse_age_vowel <- nurse_social_segment %>%
  select(Age_category, Gender, vowel)%>%
  table()/246*100
#head(nurse_age_vowel)

ggplot(nurse_social_segment, aes(x = vowel, fill = Gender)) +
  geom_bar() +
  facet_wrap(vars(Age_category), ncol = 3) +
  labs(x = "NURSE vowels", y = "Freq of NURSE vowels")+
   ggtitle("Frequencies of NURSE vowel production by social group")
```

By gender, `NURSE vowel` was produced as `[ɑ]` by 8 females and 15 males, as [ɒ] by 7 females and 11 males. Males produced the vowel as [æ] 38 times more frequently than females, while females produced "ɛ" 11 times more than males (18). Furthermore, "ɜ" was primarily produced by females, with 5 instances compared to only one among male speaker. These findings suggest that there may be gender-related variations in the production of NURSE vowels in the sample data.

```{r}
model_nurse_age_det <- glmer(nurse_determinant ~ 1 + phonological_environment + Gender + Age_category + (1|word), data = nurse_social_segment, family = poisson)
summary(model_nurse_age_det)
model_nurse_age_det %>% tab_model()
#model_nurse_age_det %>% plot_model()
plot_model(model_nurse_age_det, type = "pred", terms = c("Gender", "Age_category", "phonological_environment"))
#cat_plot(poisson.model2, pred = tension, modx = wool, geom = "line", plot.points = TRUE)
#library(jtools)
#cat_plot(model_nurse_age_det, pred = word, modx = nurse_determinant)
```

```{r}
#ggplot(nurse_social_segment, aes(x = vowel, fill = word)) +
 # geom_bar() +
  #facet_wrap(vars(Age_category), ncol = 3) +
  #labs(x = "NURSE vowels", y = "Freq of NURSE vowels")+
   #ggtitle("Frequencies of NURSE vowel production by social group")
```


```{r}
Nurse_segment <- lmer(nurse_determinant ~ 1 + phonological_environment*Gender + phonological_environment*Age_category + (1|word),  data = nurse_social_segment)
Nurse_segment %>% summary()
Nurse_segment %>% tab_model()
plot_model(Nurse_segment, type = "pred", terms = c("Age_category", "phonological_environment", "Gender"))
```

We projected the phonological environment, gender and age category would determine NURSE vowel production. The intercept, 0.61, represents the determinant's baseline occurrence rate. The phonological context, particularly when not coming after a consonant, has a coefficient of 0.28, indicating a minor increase in occurrence rate, albeit statistically insignificant (p = 0.254). Gender and age category (younger) have correlations of -0.08 and 0.14, showing no significant effect on the prediction (p > 0.05). The interaction factors between phonological environment and gender, as well as phonological environment and age category, are statistically insignificant (p > 0.05). The random effects study shows a moderate level of variability in the determinant between terms, with an intraclass correlation value (ICC) of 0.31. Overall, the model accounts for a small amount of the determinant's variation, with a marginal R-squared value of 0.027 and a conditional R-squared value of 0.324.



### Duration patterns of NURSE vowel production in NE by age and gender

```{r}
nurse_social_segment %>%
  group_by(Age_category) %>%
  summarize(age_vdur = mean(duration.cen))

age_nurse_vowel <- lmer(duration.cen ~ 1 + Age_category*duration.cen + duration.cen*Gender + (1|vowel),  nurse_social_segment) 
age_nurse_vowel %>% summary()
age_nurse_vowel %>% tab_model()

#plot_model(age_nurse_vowel, type = "pred", terms = c("Age_category", "Gender"))
```
Hooray! Younger Nigerian English speakers have short duration for NURSE vowel than the adult, however, the model
The next mixed model examined the duration effect. The intercept of 0.02 represents the baseline duration for the NURSE vowel when every other factor is zero. Age group (younger) and gender (male) have negative correlations of -0.02, indicating shorter durations for younger speakers and males compared to their peers. The interaction factors for duration and age group, as well as duration and gender, had coefficients of 0.56 and 0.93, respectively. This shows that duration has different effects on different age groups and genders, with younger people and male speakers experiencing bigger changes in NURSE vowel duration. However, the random effects for word show a minimal level of variation in duration between occurrences, with an intraclass correlation coefficient (ICC) of 0.04. This model also reveals the percentage of the variance in duration, with a marginal R-squared value of 0.905 and a conditional R-squared value of 0.909.

#### Visualization
```{r}
ggplot(nurse_social_segment, aes(x = vowel, y = duration.cen, color = Gender)) +
  geom_boxplot() +
  facet_wrap(vars(Age_category), ncol = 3) +
  labs(x = "NURSE vowels", y = "Duration of NURSE vowels")+
   ggtitle("Duration patterns of NURSE vowel production by social group")
```


## NURSE vowel production by `ethnicity` (Geographical location of the speakers)

```{r}
#NURSE vowel production by ethnicity
nurse_freq_ethnicity <- table(nurse_social_segment$vowel, nurse_social_segment$Ethnicity)
head(nurse_freq_ethnicity)
#percentage
nurse_vowel_ethnicity <- nurse_social_segment %>%
  select(Ethnicity, nurse_vowel_space)%>%
  table()/246*100
head(nurse_vowel_ethnicity)

#visualization
ggplot(nurse_social_segment, aes(x = nurse_vowel_space, fill = Ethnicity)) +
  geom_bar() +
  labs(x = "NURSE vowels", y = "Freq of NURSE vowels")+
   ggtitle("Frequencies of NURSE vowel production by ethnicity")
```

```{r}
age_nurse_ethnicity <- lmer(duration.cen ~ 1 + Ethnicity + (1|word),  nurse_social_segment) 
age_nurse_ethnicity %>% summary()
age_nurse_ethnicity %>% tab_model()
```

This linear mixed model investigated how ethnicity influenced the NURSE vowel duration. The intercept for baseline length is estimated at 0.054, with a standard error of 0.0077 and a highly significant t-value of 6.981 (p < 0.001). The ethnicity variables, notably Igbo and Yoruba, have estimates of -0.003762 and -0.029069, respectively. Yoruba ethnicity's negative coefficient (-0.029069) indicates a shorter duration than the Hausa (reference group), with a significant t-value of -3.629 (p < 0.001). However, the -0.003762 coefficient for Igbo ethnicity is not statistically significant (p = 0.736). The random effects study finds that duration varies across the ethnicity (SD = 0.03) and the word as a standard deviation of 0.04.


```{r}
ggplot(nurse_social_segment, aes(x = Ethnicity, y = duration.cen, color = Gender)) +
  geom_boxplot() +
  facet_wrap(vars(nurse_vowel_space), ncol = 3) +
  labs(x = "NURSE vowels", y = "Duration of NURSE vowels")+
   ggtitle("Duration patterns of NURSE vowel production by Ethnicity")
```

## NURSE vowel determinant by ethinicity

```{r}
nurse_ethn_det <- glmer(nurse_determinant ~ 1 + Ethnicity*phonological_environment + (1|word), data = nurse_social_segment, family = poisson)
summary(nurse_ethn_det)
nurse_ethn_det %>% tab_model()
sjPlot::tab_model(nurse_ethn_det)
plot_model(nurse_ethn_det, type = "pred", terms = c("Ethnicity"))
```
## Formant trajectories by ethnicity
```{r}
#join these two dataframe by vowel
nurse_social_segment
my_nurse_vowel_formant
ethni_formant <- merge(nurse_social_segment, my_nurse_vowel_formant, by = 'vowel')
```
```{r}
ggplot(ethni_formant, aes(x = timepoint, y = formant_bark_norms, group = id.y, color = formant)) +
  geom_line() + 
  facet_wrap(~Ethnicity, nrow = 2)+
  ggtitle("Formant trajectories for NURSE vowels across ethnicity")+
  theme_classic()
```
```{r}
#use dummy code for timepoint to identify how much formant trajectory varies across ethnicity
ethni_formant <- ethni_formant%>% 
  mutate(timepoint.num = as.numeric(timepoint))

formant_by_ethnicity <- lmer(formant_bark_norms ~ 1 + Ethnicity + (1|vowel) + (1|id.y), data = ethni_formant)
summary(formant_by_ethnicity)

#the model did not converge
```



# Session info
```{r}
sessionInfo()
```